### 2015-05-13 corral.tex
update

\medskip\noindent\textbf{Sound Partial-Order Reduction }
%
By default, \corral instruments a \texttt{yield} statement (which denotes a potential nondeterministic context-switch) before each \emph{visible operation} (i.e., synchronization operations and shared memory accesses) of every thread. It then leverages sequentialization to explore all possible thread interleavings up to a pre-defined bound.

Our approach to accelerating \corral is simple and yet effective: we simplify
the sequentialization that \corral performs by instrumenting \texttt{yield}
statements only at shared memory accesses that \whoop reported as potentially
racy.\footnote{In the presence of data races and relaxed memory, even considering all interleavings of shared memory accesses may be insufficient to find all bugs.} This effectively acts as a partial-order reduction that can significantly
reduce the number of interleavings that \corral needs to explore, and thus
increase performance and scalability.

\ZRComment{Could this be cleaned up a bit? In particular, do we need to talk
about entry point pairs, or could we prove it just for any pair?}\PDComment{Lets also see what Ally thinks about this part.}
We now show that our partial-order reduction is sound: it does not miss any more
bugs than what \corral would miss even without taking advantage of the \whoop
race-related information. Let \corral run an $n$ amount of entry point pairs
with yields in all visible operations. Let this configuration be known as $C$.
Each entry point pair is running in a separate \corral process. \corral will
miss all bugs that are triggered by more than 3 bugs running concurrently and
all bugs that are triggered by investigating a larger than $r$ recursion bound
and larger than $k$ context-switch bound. Let \corral then run the same $n$
amount of pairs, but this time with all \texttt{yield} statements in non-racy
shared memory accesses that are removed according to information from \whoop. Let this configuration be
known as $W+C$. Because these shared memory accesses cannot race, performing a
\texttt{yield} will not introduce any new program behaviors. Thus, \corral will
find the same bugs using the configuration $W+C$, as using the configuration
$C$.

@@

\yieldcoarse instruments \texttt{yield} statement in a binary fashion: if \whoop
finds that a pair of entry points cannot race, then it will only instrument
context-switches in synchronization points of this pair; else if \whoop finds
that a pair of entry points can race, then it will instrument context-switches
in all visible operations of this pair.

>>>>

\medskip\noindent\textbf{Sound Partial-Order Reduction }
%
By default, \corral instruments a \texttt{yield} statement (which denotes a potential nondeterministic context-switch) before every shared memory access of each thread. It then leverages sequentialization to explore all possible thread interleavings up to a pre-defined bound. Our approach to accelerating \corral is simple and yet effective: we simplify the sequentialization that \corral performs by instrumenting \texttt{yield} statements only at shared memory accesses that \whoop reported as potentially racy.\footnote{In the presence of data races and relaxed memory, even considering all interleavings of shared memory accesses may be insufficient to find all bugs.} This effectively acts as a partial-order reduction that can significantly reduce the number of interleavings that \corral needs to explore, and thus increase performance and scalability.

\ZRComment{Could this be cleaned up a bit? In particular, do we need to talk
about entry point pairs, or could we prove it just for any pair?}

We now show that our partial-order reduction is sound: it does not miss any more bugs than what \corral would miss even without taking advantage of the \whoop race-related information. Let \corral run an $n$ amount of entry point pairs with yields in all \emph{visible} operations (synchronization points and shared memory accesses). Let this configuration be known as $C$. Each entry point pair is running in a separate \corral process. \corral will miss all bugs that are triggered by more than 3 threads running concurrently and all bugs that are triggered by investigating a larger than $r$ recursion bound and larger than $k$ context-switch bound. Let \corral then run the same $n$ amount of pairs, but this time \whoop removes the \texttt{yield} statements in all non-racy shared memory accesses. Let this configuration be known as $W+C$. Because these shared memory accesses cannot race, performing a \texttt{yield} will not introduce any new program behaviors. Thus, \corral will find the same bugs using the configuration $W+C$, as using the configuration $C$.

@@

\yieldcoarse instruments \texttt{yield} statement in a binary fashion: if \whoop
finds that a pair of entry points cannot race, then it will only instrument
context-switches in synchronization points (i.e.\ locks and unlocks) of this pair; else if \whoop finds
that a pair of entry points can race, then it will instrument context-switches
in all visible operations of this pair.

### 2015-05-13 evaluation.tex
update

\noindent
\textbf{Experimental Setup}\xspace\xspace We performed all experiments on a 3.40GHz Intel Core i7-2600 CPU with 16GB RAM running Ubuntu Linux 12.04.5 LTS, LLVM 3.5, SMACK 1.5, Z3~4.3.2, Boogie rev. 4192 and \corral rev. 534. We also used Mono~4.1.0 to run Boogie and \corral.

\noindent
\textbf{Benchmarks}\xspace\xspace We evaluate our methodology against \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel. We chose non-trivial drivers from several categories: block; char; ethernet; near field communication (nfc); universal serial bus (usb); and watchdog. We had to manually model the environment for these drivers, a process that required approximately two months of work.

\noindent
\textbf{Race-Checking}\xspace\xspace Table~\ref{tab:stats} presents statistics for all our benchmarks: lines of code (LoC); number of entry point pairs (\#Pairs); number of SMACK memory regions (\#MRs); number of potentially racy pairs identified by \whoop (\#Racy Pairs); number of potentially racy memory regions reported by \whoop (\#Racy MRs); and number of races discovered by \corral using a context-switch bound of 2 (\#Races Found). We did not discover new races with larger context-switch bounds. This might mean that these drivers only have races that manifest with at least one or two context-switches, or that \corral hit its bounds before discovering a deeper race.

We can see that \whoop reports more races than \corral. This is expected, since \whoop employs an over-approximating shared state abstraction to conservatively model the effects of additional threads when analysing an entry point pair, and because lockset analysis is inherently imprecise; both factors can lead to false alarm race reports.  On the other hand, \corral is precise, but can miss races because only a limited number of context switches are considered.  Another issue with \corral is loop coverage due to unsound loop unrolling. To tackle this, we enable the built-in loop over-approximation described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice. Finally, when we apply \corral to a pair of entry points, we are just checking the specific pair and are not accounting for the effect of other threads (see Section~\ref{corral}); this can also cause \corral to miss some races.

@@

\noindent
\textbf{Granularity of Context-Switches}\xspace\xspace Table~\ref{tab:results} presents runtime results from using \whoop, \corral and \whoop + \corral to analyze our benchmarks. All reported times, in seconds, are averages over three runs. \ADComment{Can we say something about variance?  Perhaps ask Nathan what best to say?} \corral was given a time budget of 10 hours (T.O. denotes a tool timeout), and we considered context-switch bounds (csb) of 2, 5 and 9. By default, \corral instruments context-switches (i.e. \texttt{yield} statements) in \ADComment{instead of ``in'' can we say either ``before'' or ``after'' (not sure which is accurate)} all visible operations (denoted \yieldall). \whoop + \corral, instead, uses two different context-switch instrumentation granularities, \yieldcoarse and \yieldmr (see Section~\ref{whoop:bugfinding}), and also uses all the optimisations discussed in Section~\ref{whoop:optimizations}.

@@

%When running \corral on its own, it can potentially report a spurious race (that \whoop does not report), because \corral is not enhanced with domain-specific information that \whoop uses to discard spurious races. Using \whoop + \corral, though, will inherently discard spurious data races (assuming a precise environmental model).

>>>>

\medskip\noindent\textbf{Experimental Setup }
%
We performed all experiments on a 3.40GHz Intel Core i7-2600 CPU with 16GB RAM running Ubuntu Linux 12.04.5 LTS, LLVM 3.5, SMACK 1.5, Z3~4.3.2, Boogie rev. 4192 and \corral rev. 534. We also used Mono~4.1.0 to run Boogie and \corral.

\medskip\noindent\textbf{Benchmarks }
%
We evaluate our methodology against \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel. We chose non-trivial drivers from several categories: block; char; ethernet; near field communication (nfc); universal serial bus (usb); and watchdog. We had to manually model the environment for these drivers, a process that required approximately two months of work.

\medskip\noindent\textbf{Race-Checking }
%
Table~\ref{tab:stats} presents statistics for all our benchmarks: lines of code (LoC); number of entry point pairs (\#Pairs); number of SMACK memory regions (\#MRs); number of potentially racy pairs identified by \whoop (\#Racy Pairs); number of potentially racy memory regions reported by \whoop (\#Racy MRs); and number of races discovered by \corral using a context-switch bound of 2 (\#Races Found). We did not discover new races with larger context-switch bounds. This might mean that these drivers only have races that manifest with at least one or two context-switches, or that \corral hit its bounds before discovering a deeper race.

We can see that \whoop reports more races than \corral. This is expected, since \whoop employs an over-approximating shared state abstraction to conservatively model the effects of additional threads when analyzing an entry point pair, and because lockset analysis is inherently imprecise; both factors can lead to false alarm race reports.  On the other hand, \corral is precise, but can miss races because only a limited number of context switches are considered.  Another issue with \corral is loop coverage due to unsound loop unrolling. To tackle this, we enable the built-in loop over-approximation described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice. Finally, when we apply \corral to a pair of entry points, we are just checking the specific pair and are not accounting for the effect of other threads (see Section~\ref{corral}); this can also cause \corral to miss some races.

@@

\medskip\noindent\textbf{Granularity of Context-Switches }
%
Table~\ref{tab:results} presents runtime results from using \whoop, \corral and \whoop + \corral to analyze our benchmarks, while Figure~\ref{fig:plot} plots the runtime speedups that \corral achieves using \whoop with the \yieldmr instrumentation. All reported times, in seconds, are averages over three runs. \corral was given a time budget of 10 hours (T.O. denotes timeout), and we considered context-switch bounds (csb) of 2, 5 and 9, respectively. By default, \corral instruments context-switches (i.e. \texttt{yield} statements) in \ADComment{instead of ``in'' can we say either ``before'' or ``after'' (not sure which is accurate)} all visible operations (denoted \yieldall). \whoop + \corral, instead, uses two different context-switch instrumentation granularities, \yieldcoarse and \yieldmr (see Section~\ref{corral}).

@@
